|

### 2024-09-06 manuscript/paperCMSACversion.Rmd
Brian Mills Updates added.

While any player on a hockey team can get into a fight, enforcers are exposed to many more fights than the typical NHL player. As a result of being exposed to additional contact, they are potentially exposed to additional head trauma beyond the amount associated with playing hockey. There are many questions about the long-term implications of this head trauma and its potential to result in negative health outcomes in the future. The ingrained nature of fighting into the game of hockey is greatly criticized, and there are many active debates about its current declining trend. Figure \ref{fig:enforcer_density} shows a decline in the proportion of players in the league who are considered enforcers (as defined by @Popkin2023) among players were born more recently. 
<!-- The question is, should the game continue to move away from contact-driven strategy in favor of younger, faster, highly-skilled players? -->

@@

Based on this matched cohort study, they conclude that there is no statistically significant difference in the mortality rates between enforcers and controls for either definition of enforcer. However, they do find that enforcers died on average about 10 years earlier than their non-enforcer counterparts and more frequently of drug overdose and suicide.

@@

Figures \ref{fig:KMEF} and \ref{fig:KMEP} show Kaplan-Meier plots for E-F and E-P.  There is substantial overlap in the K-M curves between the enforcer and non-enforcers groups when enforcers is defined using the EF criteria. However, the enforcer group when defined by the EP criteria exhibits evidence of lower survival probabilities starting around age 50 relative to the non-enforcer group, though it is important to remember that these two groups differ substantially in age, height, and weight.

In order to account for these differences between the groups, Cox proportional hazards models (@CoxPH) were fit: one for each definition of enforcer defined in @Popkin2023 (i.e. E-F and E-P) with time until death as the outcome variable.  In addition, the indicators for each definition of enforcer, games played (GP), height (Ht), weight (Wt), year of birth (DOB), and position indicators (Center (C), Wing (W), or Defender (D)) were added to each model as control variables.

>>>>

The use of violence to enforce social norms in the sport, of course, also has the potential to increase injury and other long-term negative effects for hockey players. While any hockey player may participate in fighting, enforcers are exposed to many more fights than the typical NHL player. As a result of being exposed to more physical contact, they are potentially exposed to additional head trauma when compared to the broader hockey player population, putting them at potentially higher risk for long-term health outcomes. The ingrained nature of fighting into the game of hockey is therefore widely criticized, which has led to declines in the use of hockey violence. For example, figure \ref{fig:enforcer_density} shows a decline in the proportion of enforcers (as defined by @Popkin2023) to non-enforcers as players get younger. This brings forth the question of whether hockey will continue to move away from contact-driven strategies in favor of younger, faster, and highly-skilled players.

<!-- The question is, should the game continue to move away from contact-driven strategy in favor of younger, faster, highly-skilled players? -->

@@

Based on their analysis (a matched cohort study) they find that there is no statistically significant difference in the mortality rates between enforcers and controls for either definition of enforcer. However, they do find that enforcers died on average about 10 years earlier than their non-enforcer counterparts. Additionally, while there was no significant difference in all-cause mortality rate, they do suggest that enforcers in their sample died more frequently of drug overdose and suicide.

@@

Figures \ref{fig:KMEF} and \ref{fig:KMEP} show Kaplan-Meier plots for EF and EP.  There is substantial overlap in the K-M curves between the enforcer and non-enforcers groups when enforcers is defined using the EF criteria. However, the enforcer group defined by the EP criteria exhibits evidence of lower survival probabilities starting around age 50 relative to the non-enforcer group. Nevertheless, it is important to remember that these two groups differ substantially in age, height, and weight that may lead both to increased probability of becoming an enforcer and an increased probability in early death.

In order to account for these differences between the groups, we estimated full Cox proportional hazard models @CoxPH: one for each definition of enforcer defined in @Popkin2023 (i.e. E-F and E-P) with time until death as the outcome variable. Further, the indicators for each definition of enforcer, games played (GP), height (Ht), weight (Wt), year of birth (DOB), and position indicators (Center (C), Wing (W), or Defender (D)) were added as control variables.
